### 2016-04-26 manuscript.Rnw
update manuscript following rotl 3.0.0 release

<<plot_taxa, fig.height=4, cache=TRUE, purl=FALSE, fig.cap="The phylogenetic tree returned by OTL for the list of model species used as an example.">>=
tree <- tol_induced_subtree(ott_ids = taxa[["ott_id"]])
plot(tree, cex = .8, label.offset = .1, no.margin = TRUE)
@

@@

When more than one tree is available for a given study, the function
\texttt{list\_trees} returns a list containing the tree identifiers for each
study. Alternatively, the function \texttt{get\_study} returns all the trees (by
default as \texttt{phylo} objects) associated with a particular study. Metadata
about the study (e.g., citation information, information about the curator for
the study, and other technical information regarding the import of this study)
can be obtained using the function \texttt{get\_study\_meta}.

@@

In recent years, R has become an essential part of the toolbox of many
researchers in evolutionary biology and ecology. R greatly facilitates the
analysis of large datasets, and allows researchers to combine methods in novel
ways because many methods for comparative analyses are implemented, and because
it is a relatively easy to use programming language. Additionally, as more data
are made available online and accessible using APIs, several packages have been
developed to interact and download these datasets directly in R, thereby
enabling direct and reproducible analyses.  Notably, the organization rOpenSci
(\url{https://ropensci.org}) has fostered a community of researchers who develop
tools and methods to facilitate the use of open data as well broaden the
adoption of open science practices in general \citep{Boettiger2015a}. For
instance, the rOpenSci-developed package \texttt{TreeBase}
\citep{Boettiger2012TB} allows users to access phylogenies stored in treeBASE
(\url{www.treebase.org}).  \texttt{rotl} contributes to this initiative, and
greatly extends the number of taxa for which phylogenetic data can be retrieved
within R, while allowing the data from OTL to be combined with other sources
easily.

@@

<<cat_map, purl=TRUE, echo=FALSE, size='small', fig.height=4.2, results='hide', message=FALSE, cache=TRUE, fig.cap='GBIF records for the species in \\textit{Lynx} included in the phylogeny associated with the study by \\citealt{Johnson2006}.'>>=
## load required packages. Versions used for the manuscript listed below.
library(rotl)     # v0.5.0.901
library(rgbif)    # v0.9.2
library(ggplot2)  # v2.1.0
library(maps)     # v3.1.0

@@

\texttt{rotl} is free, open source, and released under a Simplified BSD
license. Stable versions are available from the CRAN repository
(\url{https://cran.r-project.org/package=rotl}), and development versions are
available from GitHub (\url{https://github.com/ropensci/rotl}). This manuscript
was built using version 0.5.0.901-rotl-ms.2
(\url{https://github.com/ropensci/rotl/tree/v0.5.0.901-rotl-ms.2}). The package
is under active development, and authors welcome bug reports or feature requests
via the GitHub repository.  The source for this manuscript is available on
GitHub (\url{https://github.com/fmichonneau/rotl-ms}).

>>>>

<<plot_taxa, fig.height=4, cache=TRUE, purl=FALSE, fig.cap="The phylogenetic tree returned by OTL for the list of model species used as an example.">>=
tree <- tol_induced_subtree(ott_ids = ott_id(taxa))
plot(tree, cex = .8, label.offset = .1, no.margin = TRUE)
@

@@

When more than one tree is available for a given study, the function
\texttt{list\_trees} returns a list containing the tree identifiers for each
study. Alternatively, the function \texttt{get\_study} returns all the trees (by
default as \texttt{phylo} objects) associated with a particular study. Metadata
about the study (e.g., citation information, information about the curator for
the study, and other technical information regarding the import of this study)
can be obtained using the functions \texttt{get\_study\_meta} and
\texttt{study\_external\_IDs}.

@@

In recent years, R has become an essential part of the toolbox of many
researchers in evolutionary biology and ecology. R greatly facilitates the
analysis of large datasets, and allows researchers to combine methods in novel
ways because many methods for comparative analyses are implemented, and because
it is a relatively easy to use programming language. Additionally, as more data
are made available online and accessible using web APIs, several packages have
been developed to interact and download these datasets directly in R, thereby
enabling direct and reproducible analyses.  Notably, the organization rOpenSci
(\url{https://ropensci.org}) has fostered a community of researchers who develop
tools and methods to facilitate the use of open data as well broaden the
adoption of open science practices in general \citep{Boettiger2015a}. For
instance, the rOpenSci-developed package \texttt{TreeBase}
\citep{Boettiger2012TB} allows users to access phylogenies stored in TreeBASE
(\url{https://treebase.org}).  \texttt{rotl} contributes to this initiative, and
greatly extends the number of taxa for which phylogenetic data can be retrieved
within R, while allowing the data from OTL to be combined with other sources
easily.

@@

<<cat_map, purl=TRUE, echo=FALSE, size='small', fig.height=4.2, results='hide', message=FALSE, cache=TRUE, fig.cap='GBIF records for the species in \\textit{Lynx} included in the phylogeny associated with the study by \\citealt{Johnson2006}.'>>=
## load required packages. Versions used for the manuscript listed below.
library(rotl)     # v3.0.0
library(rgbif)    # v0.9.3
library(ggplot2)  # v2.1.0
library(maps)     # v3.1.0

@@

\texttt{rotl} is free, open source, and released under a Simplified BSD
license. Stable versions are available from the CRAN repository
(\url{https://cran.r-project.org/package=rotl}), and development versions are
available from GitHub (\url{https://github.com/ropensci/rotl}). This manuscript
was built using \texttt{rotl} 3.0.0
(\url{https://github.com/ropensci/rotl/tree/v3.0.0}). The package
is under active development, and authors welcome bug reports or feature requests
via the GitHub repository.  The source for this manuscript is available on
GitHub (\url{https://github.com/fmichonneau/rotl-ms}).
